Online (Bandit) Policy Selection for EASY-Backfilling

### 2017-03-31 paper/src/paper.tex
include Eric's modifications.

\usepackage{listings}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[scientific-notation=true]{siunitx}

@@

Several works use \textbf{this method} (LAQUELLE?) in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that
better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.
However, these works do not address the duality between the cumulative and
maximal scheduling costs, as mentioned in~\cite{learningruntimes}.
While these previous works intend to estimate uncertain parameters, we
consider here a more
pragmatic approach, which directly learn a good scheduling policy from a
given policy space.

@@

We present here the strategies we have retained for selecting a policy. In the remainder, we will refer to the period during which a selected policy is applied as the \textit{policy period} and will denote the length of this period as $\Delta$. The time interval is thus divided into periods of equal length: $\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy period; all periods have length $\Delta$. At the beginning of each period, from the first period $\Delta_0$ to the current period $\Delta_T$, a new policy is selected.

We assume here that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods. This assumption is validated in the study presented in~\ref{???}. It further entails that the behavior of a policy on the previous periods reflects its behavior on the current one, so that the selection of a policy can be based on its behavior on previous periods. However, there usually is an important variability on the cost metric between different periods, for any policy. This is illustrated in Figure~\ref{fig:variability} where \textcolor{orange}{???}. When the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation.

@@

\subsection{Online policy selection with exact simulation}
\label{sub:feedback}

Several simulators have been developed for "playing" reordering policies on a given set of jobs. Such simulations are interesting inasmuch as they provide an estimate of the cost of a given policy on a set of jobs. \textcolor{orange}{TO BE COMPLETED}

Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 10$) one reordering policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$ during the period $\Delta_t$ is defined as:
%
\begin{equation}
\label{eq:cost-exact}
w_{\mbox{exact}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \mbox{Wait}_j^{P_i}
\end{equation}
%
where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of the waiting time for job $j$ according to policy $P_i$. The estimation of the cost of a policy over all the periods preceding the current period can then be defined as:
%
\begin{equation}
\label{eq:tot-cost-exact}
w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{exact}}(t;P_i)
\end{equation}
%
$\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent history (\textit{i.e.} recent periods). Note that the above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.

In the context of this study, as detailed in Section~\ref{sec:experiments}, we rely on metrics averaged over several execution traces in order to obtain reliable estimate of the behavior of different policies and policy selection strategies. Such traces are typically obtained by simulation. Using the same simulator for generating the traces and estimating the cost as defined in Eqs \ref{eq:cost-exact} and \ref{eq:tot-cost-exact} would however be too optimistic and would represent an upper bound on what can be achieved by a selection strategy based on simulation.

In order to have a more realistic estimate of the behavior of a selection strategy based on simulation, we introduce noise in the simulator, as described below.

\subsection{Online policy selection with noisy simulation}
\label{sub:noisy}

In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling, either dow or up, the waiting time of jobs obtained by simulation:
%
\begin{equation}
\label{eq:cost-noisy}
w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho \mbox{Wait}_j^{P_i}
\end{equation}
%

\subsection{Bandit policy selection}
\label{sub:bandit}
When we have only bandit feedback of the accurate simulator.
We use the epsilon-greedy bandit\ref{alg:epsg}.

\begin{algorithm}[h]
  \caption{Epsilon-Greedy Bandit policy}
  \begin{algorithmic}[1]
    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
    \renewcommand{\algorithmicensure}{\textbf{Output:}}
    \REQUIRE $0 < \epsilon < 1$, $K>1$
    \FOR {$t = 1, 2, \ldots$}
    \STATE Let $i_t$ be the reordering heuristic with the highest current
    average reward.
    \STATE With probability $1-\epsilon$ use $i_t$ and with probability
    $\epsilon$ use a random reordering heuristic.
    \ENDFOR
  \end{algorithmic}
  \label{alg:epsg}
\end{algorithm}

>>>>

\usepackage{listings}
\usepackage{dirtree}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{float}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bbm}
\usepackage[scientific-notation=true]{siunitx}

@@

Several works use \textbf{this method} (LAQUELLE?) in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that
better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.
However, these works do not address the duality between the cumulative and
maximal scheduling costs, as mentionned in~\cite{learningruntimes}.

A recent work~\cite{mlmem} proposes to help
users specify better memory requirements by applying supervised learning
techniques to job metadata.

While these previous works intend to estimate uncertain parameters, they lack
in validating the impact on the system. We consider in this paper a more
pragmatic approach, which is to directly learn a scheduling policy from a given
seach space.

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[scale=0.6]{figures/variability.pdf}
  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (see Subsection~\ref{sub:traces}) for 7 different policies.}
  \label{fig:mosn}
\end{figure*}

We present here the strategies we have retained for selecting a policy. We will refer to the period during which a selected policy is applied as the \textit{policy period} and will denote the length of this period as $\Delta$. The time interval is thus divided into periods of equal lengths ($\Delta$): $\Delta_0, \cdots, \Delta_T$, where $T$ is the index of the current policy period. A new policy is selected at the beginning of each period and applied during the whole period.

We further assume that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods. This assumption is validated in the study presented in~\ref{???}. It further entails that the behavior of a policy on the previous periods reflects its behavior on the current one, so that the selection of a policy can be based on its behavior on previous periods. However, there may be some variability between different periods for certain cost metrics. This is illustrated in Figure~\ref{fig:mosn} that displays the average waiting time for a trace using weekly periods encompassing one year. As one can note, the average waiting time varies a lot from one week to the other, for all the 7 policies considered. This indicates that when when the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation. We will come back to this issue below.

@@

\subsection{Policy selection with exact simulation}
\label{sub:feedback}

Several simulators have been developed for "playing" reordering policies on a given set of jobs. We rely in this study on a lightweight simulator (see Section~\ref{sec:experiments}) that can efficiently simulate different policies. Such simulators are interesting inasmuch as they provide an estimate of the cost of a given policy on a set of jobs, as described below.

Let $l(\Delta_t)$ denote the number of jobs \textit{submitted} during the period $\Delta_t$ ($0 \le t < T$), and $P_i$ ($1 \le i \le 10$) one reordering policy (defined in Section~\ref{subsec:policies}). The cost of policy $P_i$ during the period $\Delta_t$ is defined as:
%
\begin{equation}
\label{eq:cost-exact}
w_{\mbox{exact}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \mbox{Wait}_j^{P_i}
\end{equation}
%
where $\mbox{Wait}_j^{P_i}$ denotes the estimate provided by the simulator of the waiting time for job $j$ according to policy $P_i$. The estimation of the cost of a policy over all the periods preceding the current period can then be defined as:
%
\begin{equation}
\label{eq:tot-cost-exact}
w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{exact}}(t;P_i)
\end{equation}
%
$\lambda \in [0,1]$ is a decay parameter that can be used to privilege recent history (\textit{i.e.} recent periods). One then selects the policy $P$ that minimizes the above cost for the period $\Delta_T$:
%
\begin{equation}
\label{eq:select-exact}
P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
\end{equation}

The above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.

Furthermore, in the context of this study and as detailed in Section~\ref{sec:experiments}, we rely on averages over several execution traces in order to obtain reliable estimate of the behavior of different policies and policy selection strategies. Such traces are typically obtained by simulation. Using the same simulator for generating the traces and estimating the cost as defined in Eqs \ref{eq:cost-exact} and \ref{eq:tot-cost-exact} would however be too optimistic and would represent an upper bound on what can be achieved by a selection strategy based on simulation. In order to have a more realistic estimate of the behavior of a selection strategy based on simulation, we introduce noise in the simulator, as described below.

\subsection{Policy selection with noisy simulation}
\label{sub:noisy}

In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling it, either down or up, for each job:
%
\begin{enumerate}
\item For each job $j$ submitted during $\Delta_t$: $\rho_j \sim U[0.85,1.15]$
\item Then:
%
\begin{equation}
\label{eq:cost-noisy}
w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho_j \, \mbox{Wait}_j^{P_i}
\end{equation}
%
\end{enumerate}
%
$\rho$ is uniformly sampled in the interval $[0.85,1.15]$, thus adding a +/-15\% noise on the waiting time estimated by the simulator. The overall cost is then defined in the same way as before, leading to:
%
\begin{equation}
\label{eq:tot-cost-noisy}
w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i)
\end{equation}
%
As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above cost is selected for the period $\Delta_T$.

Instead of introducing noise in the output of the simulator when selecting the policy, one could have introduced it when generating the execution traces. This would however have prevented us from comparing the exact and noisy selection methods, the first one representing an upper bound on the second one. As we will see in Section~\ref{sec:experiments}, there is little difference between the two selection methods, which is an important result of our study.

Lastly, as before, the length of the period has no impact here as the cost metric corresponds to the cumulative simulated waiting time for each policy.

\subsection{Policy selection with $\epsilon$-greedy exploration}
\label{sub:bandit}

The previous strategies estimate the cost of all policies over all previous periods; the best policy according to this cost is then selected for the current period, the costs of all policies being updated for the next period. This is interesting as one maintains a complete knowledge of all policies over time. However, this requires computing many estimates at each period (as many as there are policies), which can be time consuming even with lightweight simulators.

We thus explore here a more efficient strategy that dispenses with estimating the cost of all policies at a given time and only updates the cost of the policy that is currently being used. This strategy makes use of the $\epsilon$-greedy exploration method, standard in reinforcement learning \cite{Watkins:1989}, to tradeoff exploitation and exploration, and relies, as before, on past estimates of the cost to select the best policy in the exploitation mode.

Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during the period $\Delta_t$. We define the cost of policy $P_i$ during the period $\Delta_t$ as:
%
\[
w_{\epsilon}(t;P_i) =
\left\{
\begin{array}{ll}
 \sum_{j=1}^{l(\Delta_t)} \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t \nonumber\\
 0 & \mbox{otherwise}
 \end{array}
\right.
\]
%
As one can note, the actual waiting time is used here, so that one dispenses with the use of a simulator for efficiency considerations. This indeed leads to faster and more accurate estimate of the cumulative waiting time, however only for the policy that is used during $\Delta_t$. This also explains the difference between the definitions of $l(\Delta_t)$: in the previous selection strategies, one relies on a simulated estimate so that jobs need not be finished during $\Delta_t$, but only submitted; in contrast, to know their actual waiting time, jobs need be finished during $\Delta_t$.

The cost over all previous period of a policy $P_i$ is then defined as:
%
\begin{equation}
\label{eq:tot-cost-greedy}
w(\rightarrow T;P_i) = \frac{1}{\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)} \sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) \lambda^{T-1-t} w_{\epsilon}(t;P_i)
\end{equation}
%
where $\mathbbm{1}(P_i,t) = 1$ if $P_i$ is the policy used during the period $\Delta_t$ and $0$ otherwise. The normalization by $\sum_{t=0}^{T-1} \mathbbm{1}(P_i,t) l(\Delta_t)$ is necessary here to ensure that policies remain comparable over time, as not normalizing would favor the policy that was first selected, even if this choice was random. This normalization however entails that the size of the policy period is important: it should not be too large so as to avoid relying on too few points for estimating the cost, and it should not be too small to avoid extreme variations between periods. As explained in Section~\ref{sec:experiments}, we rely in this study on periods of one day and one week.

The selection of the policy to be used during the current time period $\Delta_T$ is then based on the standard $\epsilon$-greedy exploration strategy:
%
\begin{itemize}
\item With probability $(1-\epsilon)$, select the policy $P$ that minimizes the cost over previous time periods (exploitation mode):
%
\begin{equation}
\label{eq:select-greedy}
P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
\end{equation}
%
\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 10$, at random (exploration mode).
\end{itemize}

Several studies propose to update $\epsilon$ over time, the exploration being less important once the estimates for the different policies are reliable \cite{Tokic:2010}. This strategy has however not been beneficial in our case. We believe this is due to the stochastic nature of the traces we use, as explained in Section~\ref{sec:experiments}.

%\begin{algorithm}[h]
%  \caption{Epsilon-Greedy Bandit policy}
%  \begin{algorithmic}[1]
%    \renewcommand{\algorithmicrequire}{\textbf{Input:}}
%    \renewcommand{\algorithmicensure}{\textbf{Output:}}
%    \REQUIRE $0 < \epsilon < 1$, $K>1$
%    \FOR {$t = 1, 2, \ldots$}
%    \STATE Let $i_t$ be the reordering heuristic with the highest current
%    average reward.
%    \STATE With probability $1-\epsilon$ use $i_t$ and with probability
%    $\epsilon$ use a random reordering heuristic.
%    \ENDFOR
%  \end{algorithmic}
%  \label{alg:epsg}
%\end{algorithm}
